---
title: "SNP_Primer_selection"
output: html_document
date: "2022-11-17"
---
This script is meant to be used after BatchPrimer3 SNP picking from the fasta created by bedtools. make sure to rename and save as a csv

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r load libraries}
library(here)

```

BatchPrimer3 only allows 500 Seqs at a time, so we have 6 batches of primer sets to import
```{r data import}
B1 <- read.csv(here("Seal_ID_SNP_panel","SNPbatch1_BP3.csv"), header = T)
B2 <- read.csv(here("Seal_ID_SNP_panel","SNPbatch2_BP3.csv"), header = T)
B3 <- read.csv(here("Seal_ID_SNP_panel","SNPbatch3_BP3.csv"), header = T)
B4 <- read.csv(here("Seal_ID_SNP_panel","SNPbatch4_BP3.csv"), header = T)
B5 <- read.csv(here("Seal_ID_SNP_panel","SNPbatch5_BP3.csv"), header = T)
B6 <- read.csv(here("Seal_ID_SNP_panel","SNPbatch6_BP3.csv"), header = T)
```


```{r}
library(tidyverse)
library(stringr)
```
We want all of the SNPs in one DF

```{r}
Primers <- rbind(B1,B2,B3,B4,B5,B6)
# so there are 1984 total SNPs to choose from

```

We need better names for the primers so we know they are harbor seals
```{r}
Primers %>%
  separate(col = Seq.ID, into = c("g","db","PrimerID"), sep = "\\|") %>%
  select(!(g:db)) -> Primers

Primers$PrimerID <- str_replace(Primers$PrimerID, "rs", "Pv")

```

We want to be sure to include the ones that were shared between the Pacific dataset and the Salish Sea seal data
```{r}
Pac_SS <- read.csv(here("SNPs_inboth_SS_Pac.csv"), header = T)

Pac_SS %>%
  mutate(Pv = "Pv") %>%
  unite("PrimerID", sequence, position, sep="_") %>%
  unite("PrimerID", Pv, PrimerID, sep = "") -> Pac_SS

# pull out the ones that are in both
keep <- inner_join(Pac_SS, Primers, by = "PrimerID")
# 22 of 28 had primers created by BatchPrimer3, so we need to keep those, and 2 for the sex ratio so 24 F/R primer sets to order are for sure. So 744 more sets to choose for 8 96well plates

Primers2 <- anti_join(Primers, keep, by = "PrimerID")

Primers2 %>%
  filter(Orientation == "FORWARD") -> Forward
Primers2 %>% 
  filter(Orientation == "REVERSE") -> Reverse
```

So for the rest we need to randomly select 744 of the 1963 remaining sets
```{r}
Forward %>%
  mutate(PrimNum = 1:n()) %>%
  mutate(direction = "F") -> Forward
Reverse %>%
  mutate(PrimNum = 1:n()) %>%
  mutate(direction = "R")-> Reverse

rbind(Forward,Reverse) -> Primers3
Primers3 %>%
  arrange(PrimNum) -> Primers3

RandV <- sample(1:1963, 744, replace = F)

Primers3 %>%
  subset(PrimNum %in% RandV) -> RandomSelect

RestPrim <- anti_join(Primers3, RandomSelect)

write.csv(RestPrim, "PrimersNotSelected.csv", row.names = F)

RandomSelect %>%
  unite("PrimerName", PrimerID, direction) -> Primers4

write.csv(Primers4, "PrimersForOrdering.csv", row.names = F)
```

Use the template from WDFW to use their excel sheet to add the adapters to the primers and then port them into the ordering sheet
